---
title: Predicting the manner and quality of Barbell Lifts using data from Fitness
  Devices
author: "Nakul Sapru"
date: "23/10/20"
output:
  pdf_document: default
  html_document: default
---



**Introduction**

The aim of this project is to build a predictive model to know how a particular user is lifting weights based on data obtained from an accelerometer.

The dataset consists of 5 classes:

Class A => The subject is lifting weights exactly according to the specifications.

Class B => Throwing the elbow to the front.

Class C => Lifting the dumbbell only halfway.

Class D => Lowering the dumbbell only halfway.

Class E =>Throwing the hips to the front.

Further details about the dataset can be found using the following URL: <http://groupware.les.inf.puc-rio.br/har>

**Getting The Data**

The file "pml-training" will be used as the training set.
The file "pml-testing" is a data set without the classes i.e. the classes will be predicted using the model.
```{r}
if(!file.exists("pml-training.csv"))
{
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "pml-training.csv", method = 'curl')
}
dataset <- read.csv("pml-training.csv", na.strings = c("NA", ""))
if(!file.exists("pml-testing.csv"))
{
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "pml-testing.csv", method = 'curl')
}
validation <- read.csv("pml-testing.csv")
```

**Data Preprocessing**

The necessary packages are being imported below.
```{r}
library(caret)
library(randomForest)
```

A suitable seed is being set below to ensure reproducibility.

```{r}
set.seed(17)
```

Once the data is partitioned, 70% of it will go to the training set and the rest will be the test set.

```{r}
inTrain = createDataPartition(y=dataset$classe, p=0.7, list=FALSE)
training = dataset[inTrain,]
testing = dataset[-inTrain,]
```

The NA entries are being eliminated below.

```{r}
naColumns = sapply(training, function(x) {sum(is.na(x))}) #Make a vector of all the columns and the number of NA entries
naColumns
columnsWithNA = names(naColumns[naColumns > 0]) #Vector with all the columns that has NA values
training = training[, !names(training) %in% columnsWithNA] #Remove those columns from the training set
names(training)
#Remove unnecessary columns (the first 7 columns)
training <- training[, !names(training) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

The same procedure is followed for the validation set.

```{r}
naColumns = sapply(validation, function(x) {sum(is.na(x))}) #Make a vector of all the columns and the number of NA entries
columnsWithNA = names(naColumns[naColumns > 0]) #Vector with all the columns that has NA values
validation = validation[, !names(validation) %in% columnsWithNA] #Remove those columns from the training set.
validation <- validation[, !names(validation) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

The same procedure is followed for the testing set.

```{r}
naColumns = sapply(testing, function(x) {sum(is.na(x))}) #Make a vector of all the columns and the number of NA entries
columnsWithNA = names(naColumns[naColumns > 0]) #Vector with all the columns that has NA values
testing = testing[, !names(testing) %in% columnsWithNA] #Remove those columns from the training set.
testing <- testing[, !names(testing) %in% c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")]
```

Next, the predictive model is built using Random Forest.

```{r acc}
model <- randomForest(classe ~ .,   data=training, ntree = 50)
predictions <- predict(model, testing)
confusionMatrix(predictions, testing$classe)
modelAcc <- confusionMatrix(predictions, testing$classe)$overall[[1]]
```

The model is `r modelAcc` accurate.

Now, the unknown classes of the validation set are predicted.

```{r}
predictions <- predict(model, validation)
predictions
```